---
title: "prosperLoanData"
author: "Ramin Mohammadi"
date: "May 9, 2017"
output: 
html_document: 
fig_caption: yes
keep_md: yes
number_sections: yes
toc: yes
---

```{r, include=FALSE, cache=TRUE}
#install.packages("tidyr")
#install.packages("memisc")


library("plyr") 
library("tidyr")
library("dplyr")
library(ggplot2)
library(gridExtra)
library(reshape2)
library(memisc)
require(knitr)
require(dplyr)
require(GGally)
require(psych)
require(Hmisc)


```


<h2>Abstract</h2>

I explore a data set about Prospect loans p2p. My goals with the study are 
understanding how this practice works and find what seems to affect the borrower
rate.


```{r echo=FALSE , message=FALSE , warning=FALSE}

setwd("~/Downloads/EDA_Project_Ramin_Mohammadi/")
pld<-read.csv("prosperLoanData.csv", stringsAsFactors = FALSE ,
              na.strings = c("","NA"))
```

<h2>Introduction</h2>

Dataset has 113937 records with 81 variables (including Date, characters, 
and numeric)

### Summary of data set

```{r echo=FALSE , message=FALSE , warning=FALSE}

theme_set(theme_minimal(17))
str(pld)
```


# **variables selection**

Following variable have been dropped as in order to make the analysis easier to 
undrestand

```{r echo=FALSE , message=FALSE , warning=FALSE}

theme_set(theme_minimal(17))
variables_to_drop<-c("ListingKey","LenderYield","EstimatedEffectiveYield"
                     ,"EstimatedLoss","EstimatedReturn"
                     ,"ProsperRating","ProsperRating","ListingCategory",
                     "Occupaation","EmploymentStatusDuration","CurrentlyInGroup"
                     ,"GroupKey","DateCreditPulled","FirstRecordedCreditLine",
                     "TotalCreditLinespast7years","OpenRevolvingAccounts"
                     ,"OpenRevolvingMonthlyPayment","PublicRecordsLast10Years"
                     ,"PublicRecordsLast12Months","RevolvingCreditBalance",
                     "TradesOpenedLast6Months","IncomeVerifiable","LoanKey",
                     "OnTimeProsperPayments","LoanFirstDefaultedCycleNumber"
                     ,"LoanMonthsSinceOrigination","LoanNumber","MemberKey",
                     "LP_NonPrincipalRecoverypayments"
                     ,"Investors")
```

```{r echo=FALSE , message=FALSE , warning=FALSE}

theme_set(theme_minimal(17))
pld<-pld[,!names(pld)%in%variables_to_drop]
```

looking throug dataset and droping column which have more than 2000 records
empty or NA

```{r echo=FALSE , message=FALSE , warning=FALSE}
na_count<-sapply(pld, function(x) sum(length(which(is.na(x)))))
pld<-pld[,-(which(na_count>20000))]
```



# Univariate ploting

```{r echo=FALSE , message=FALSE , warning=FALSE}

ggplot(aes(x =LoanStatus ) , data=pld) + geom_bar() + 
theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

As we can see we are having mutiple 'past due' based on different day values , 
for simplicity and making our plot to be easier to undrestand i convert all 
values whihc start with "Past" to "Past Due" 


```{r echo=FALSE , message=FALSE , warning=FALSE}

theme_set(theme_minimal(17))

index<-grep("Past" , pld$LoanStatus)
pld[index,4]<-"Past Due"
```

Also we can see Cancelled status is only happening 5 times out of 113932 records,
so i remove those records


```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))

n<-which(pld$LoanStatus == "Cancelled")
pld<-pld[-n,]

ggplot(aes(x =LoanStatus ) , data=pld) + geom_bar() + theme(axis.text.x = 
element_text(angle = 90, hjust = 1))
```

majority of loans in our datase are in current status then second majority is 
completed however we can see we have final payment in progress which are not much 
so i am removing the cases which are final payments


```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))
index<-which(pld$LoanStatus == "FinalPaymentInProgress")
pld<-pld[-index,]
table(pld$LoanStatus)
```

We can see our plot is more undrestandable 

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
months_list <- c('Jan','Feb','Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep','Oct',
'Nov', 'Dec')

pld$LoanOriginationDate <- as.Date(pld$LoanOriginationDate , format = "%Y-%m-%d", )
pld$LoanOriginationDate.month = format(pld$LoanOriginationDate , "%b")
pld$LoanOriginationDate.year = format(pld$LoanOriginationDate , "%Y")

ggplot(aes(x = LoanOriginationDate.month) , data = pld)  + 
geom_histogram(stat = "count")
ggplot(aes(x = LoanOriginationDate.year) , data = pld)  + 
geom_histogram(stat = "count")


```

Most of the loans happend on 2013 and also highest amount of loan request are 
first month of the year and october , we can check for reason in multivariate 
analysis to see what is the reason behind 


```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))

ggplot(aes(x =EmploymentStatus ) , data=pld) + geom_bar() +
theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

mixing "not available", "other" and "NA" values for Employment Status and change
them to "Not Available", Also we can see we are having a value as Employed which
can contain (Fulltime or PartTime) so for simplicity i mix them together


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

n<-which(pld$EmploymentStatus %in% c("Not available" , "Other"))
pld$EmploymentStatus[n]<-"Not Available"


n<-which(pld$EmploymentStatus %in% c("Employed" , "Full-time" , "Part-time"))
pld$EmploymentStatus[n]<-"Employed"

n<-which(is.na(pld$EmploymentStatus))
pld<-pld[-n,]
ggplot(aes(x =EmploymentStatus ) , data=subset(pld , !is.na(EmploymentStatus)))+
geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Majority of borrowers are Employed, however we have 835 case which dont have any 
information regarding their employment status and assumin that we are dealing with 
loan information means the information is missing so, i remove cases without
employment status information

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
index<-which(pld$EmploymentStatus == 'Not Available')
pld<-pld[-index,]
table(pld$EmploymentStatus)
```

We can see we have more Employed than other values which is expecting, however 
Not available is not quite clear as for getting loan you need to provide that 
information


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
ggplot(aes(x =Term ) , data=pld) + geom_bar(aes(fill = EmploymentStatus))

```

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
table(pld$Term)
```

We can see most of the Loans having 36 months terms which also expecting as having
higher terms will require paying more intrest and having lower term means monthly
payment should be more , also we can see retired class are having 36 months loans
mostly


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = log(StatedMonthlyIncome) ) ) + 
               geom_boxplot(alpha = 0.2, color = 'red' ,size =1) ,
             ggplot(pld, aes( x   = log(StatedMonthlyIncome)  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
summary(pld$StatedMonthlyIncome)
```

at this point there are 1393 case which have income of 0, my assumption is either
there has been a mistake or input is wrong as bank will not provide you a loan with
income of 0 , so for my work i get ride of records with income of 0 , ALSO i am
intrested in monthly incomes less than equal to 10000 as more than 10000 is not 
likly to happen

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
index <-which(pld$StatedMonthlyIncome <= 500 | pld$StatedMonthlyIncome > 10000)
pld<-pld[-index,]

```

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = StatedMonthlyIncome ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = StatedMonthlyIncome  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

we can see median is 4538 and we have majority of data within first quartile and third quartile (3317,6250)

```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = MonthlyLoanPayment ) ) + 
               geom_boxplot(alpha = 0.2, color = 'red' ,size =1) ,
             ggplot(pld, aes( x   = MonthlyLoanPayment  ) ) +
               geom_histogram(bins=30 ),ncol=2)

```

```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))
summary(pld$MonthlyLoanPayment)
```

looking at summary table we can see monthly loan payment of 0 which indicates of
wrong data, as we can not have monthly payment of 0, i will remove records with 
value of 0 , also we only have 473 case with monthly payment of greater than 700$
, i removed those case as well as they are outlier to me.

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
index <-which(pld$MonthlyLoanPayment <=50 | pld$MonthlyLoanPayment > 700)
pld<-pld[-index,]
```

```{r echo=FALSE , message=FALSE , warning=FALSE}
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = MonthlyLoanPayment ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = MonthlyLoanPayment  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

monthly loan payment has distribution which is right skewed 


```{r echo=FALSE , message=FALSE , warning=FALSE}  
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = CreditScoreRangeLower ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = CreditScoreRangeLower  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

```{r echo=FALSE , message=FALSE , warning=FALSE}  
summary(pld$CreditScoreRangeLower)
```

lower credite score range has ouliers based on box-plot , i consider to eliminate
records with lower credite range less than 550 and higher than 800


```{r echo=FALSE , message=FALSE , warning=FALSE}  

grid.arrange(ggplot(pld, aes( x = 1, y = CreditScoreRangeUpper ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = CreditScoreRangeUpper  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

```{r echo=FALSE , message=FALSE , warning=FALSE}  
summary(pld$CreditScoreRangeLower)

```

i have consider same condition for Upper range credit, so i rmoved credite range 
less than 550 and higher than 800



```{r echo=FALSE , message=FALSE , warning=FALSE}  
index <-which(pld$CreditScoreRangeUpper <=550 | pld$CreditScoreRangeUpper > 800 |
pld$CreditScoreRangeLower > 800 | pld$CreditScoreRangeLower <= 550)
pld<-pld[-index,]
```

```{r echo=FALSE , message=FALSE , warning=FALSE}  


grid.arrange(ggplot(pld, aes( x = 1, y = CreditScoreRangeLower ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = CreditScoreRangeLower  ) ) +
               geom_histogram(bins=30 ),
               ggplot(pld, aes( x = 1, y = CreditScoreRangeUpper ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = CreditScoreRangeUpper  ) ) +
               geom_histogram(bins=30 ),ncol=2)
```

we can see both Upper and Lowe rcredit range having similar distribution however,
they are having slightly different quartiles. later i am creating a new variable 
as average credit score and remove Upper and Lower scores.


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = LP_CustomerPayments ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = LP_CustomerPayments  ) ) +
               geom_histogram(bins=30 ),ncol=2)

```

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
summary(pld$LP_CustomerPayments)
```

we can see here that on minimum value we are having negative value which indicate 
of either wrong input or wrong value, i will remove all records with values less
than 0 and higher than 12000.

```{r echo=FALSE , message=FALSE , warning=FALSE}
index<-which(pld$LP_CustomerPayments <0 | pld$LP_CustomerPayments > 12000 )
pld<-pld[-index,]

```


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

grid.arrange(ggplot(pld, aes( x = 1, y = LP_CustomerPayments ) ) + 
               geom_boxplot(alpha = 0.2, color = 'blue' ,size =1) ,
             ggplot(pld, aes( x   = LP_CustomerPayments  ) ) +
               geom_histogram(bins=30 ),ncol=2)

```

distribution is right skewed with median as 2338 , and we can see majority have 
payment less than 4702 which is 3rd quartil.


# Bivariate Analysis


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(x =log(StatedMonthlyIncome) , y=log(AmountDelinquent))
, data=pld) + geom_point()
```

it seems that people with higher monthly incomes are more delinquent, i will 
add third variable to color the plot based on income Ranege


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(x =log(StatedMonthlyIncome) , y=log(AmountDelinquent) ,color = IncomeRange )
, data=pld) + geom_point()
```

we can see chance of being delinquent incerases as income goes higher and when 
incomes reaches to 100K slightly decrase.

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(x =InquiriesLast6Months , y=BankcardUtilization)
, data=pld) + geom_point(alpha = 0.01) + geom_jitter()

```

i have not seen much relation between bank card utilization and inquires within 
last 6 month, however it seems higher inquires tends to have abit less utilization,
so i am looking in more details and create buckets for number of inquiries

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

pld$bucket.InquiriesLast6Months <- cut(pld$InquiriesLast6Months, c(0.000,   
1.000 ,  1.435 ,  2.000 ,105.000 ))

ggplot(aes(bucket.InquiriesLast6Months, 
            BankcardUtilization, fill = EmploymentStatus), 
        data = subset(pld, !is.na(bucket.InquiriesLast6Months)) ) +
   geom_boxplot( alpha = .5)
```

intresting fact i have found is self-employes tends to always having lower median
than other people.


# Multivariate analysis


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

data<-pld[,c("BorrowerRate" ,"LoanStatus","LP_CustomerPayments","TotalInquiries"
,"InquiriesLast6Months","LoanOriginationDate.month","LoanOriginationDate.year",
"DebtToIncomeRatio","OpenCreditLines", "LP_CustomerPrincipalPayments",
"CreditScoreRangeLower","CreditScoreRangeUpper","MonthlyLoanPayment",
"StatedMonthlyIncome","LoanOriginalAmount","Term","EmploymentStatus",
"AmountDelinquent","bucket.InquiriesLast6Months","BankcardUtilization",
"IncomeRange")]


```

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

data$average_credit_score<- (data$CreditScoreRangeLower + data$CreditScoreRangeUpper)/2
data<-data[,-c(11,12)]

nums<-sapply(data , is.numeric)
data_for_correlation<-data[,nums]
```
i have created a new variable for average of credit scores ( Lower and Upper) and
removed upper and lower range variables

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
summary(data$average_credit_score )

```


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

pairs.panels(data_for_correlation[,1:4])
pairs.panels(data_for_correlation[,5:9])
pairs.panels(data_for_correlation[,10:14])
```

also for simplicity and due to having several columns it might be easier also to 
look at correlation values only 

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

cor(data_for_correlation  , use = "complete.obs" )

```

i am intrested in correlations higher than 0.3 as they seems good and as the 
correlation get closer to 1 it will be better, in this analysis i am looking at
Borrower rate and find out variables which are correlated to it and can help,
we can see Loan original amount, average credit score having highest correlation
with borrower rate.

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(x = LoanOriginalAmount , y= BorrowerRate) , data = data)+
geom_point() + geom_smooth(method = "lm")

```

first look we can see ther seems to be a relation and having higher loan amount
tends to getting lower rate, however i will go forward and create buckets for
original amount and re create the plot

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
data$bucket.LoanOriginalAmount <- cut(data$LoanOriginalAmount, c(1000,   4000 ,
6000 ,  8000 ,32000 ))

ggplot(aes(bucket.LoanOriginalAmount, 
            BorrowerRate, fill = EmploymentStatus), 
        data = subset(data, !is.na(bucket.LoanOriginalAmount)) ) +
   geom_boxplot( alpha = .5)

```

we can see as the loan amount increases, not employed customer tends to have 
higher borrower rate in compate to others, also it seems Retired customer
seems to get lower rate in compare to other groups

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

data$average_credit_score.bucket <- cut(data$average_credit_score, c(560,   669.5 ,
689.5 ,  729.5 ,800 ))

grid.arrange(ggplot(aes(y = BorrowerRate , x= average_credit_score , 
                       colour = LoanOriginationDate.year )
            , data = data)+ geom_point(alpha=0.1) + geom_jitter(),
  ggplot(aes(average_credit_score.bucket, 
            BorrowerRate, fill = LoanOriginationDate.year), 
            data = subset(data, !is.na(average_credit_score.bucket)) ) +
            geom_boxplot( alpha = .5)
            )

```

looking at the plot first i was thinking there is nothing intresting about this 
plot , however when adding parameter loan origin year for colour we can see lighter
colours which are belonging to years 2005-2008 have les borrowerRate and they had 
lower average credit score in compare to recent years ( 2008 - 2014) which we can
see avergae credit scores are above 600 and borrower rate also increases.

also defining a variable as average credit card buckets which has 5 buckets as,
first quartile, median, 3rd quartile and maximum and ploting the box plots we can
see during years 2007-2014 there is a curve which means we have an increase 
in borrower rate. we cans see year 2010 for each credit bucket is the pick and then
we have a drastical decrease in borrower rate which 2014 is the lowest, and 
having higher credit score will gives lowest borrower rate.

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(y = BorrowerRate , 
           x= factor(average_credit_score) )
           , data = data)+
  geom_boxplot(alpha = 0.1)

```
by looking at the boxes and medians we can see having higher credit score cause 
to get lower borrower rate 



```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))

ggplot(aes(x =  LoanOriginalAmount , y=MonthlyLoanPayment) , data = data)+
geom_point(alpha = 0.2) + geom_smooth(method = "lm") + geom_jitter()

```

looking at the plot there seems to be intresting linear relation in upper portion 
of the data, we can see the upper line has a monthly loan payment > 250 and loan 
original amount less than 10000 so i am subseting the data



```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
index<-which(data$MonthlyLoanPayment > 250 & data$LoanOriginalAmount <= 10000)
data<-data[-index,]

ggplot(aes(x =  LoanOriginalAmount , y=MonthlyLoanPayment , colour = 
average_credit_score) , data = data)+
geom_point(alpha = 0.2) + geom_smooth(method = "lm") + geom_jitter()

```

i have tried average_credit_score , Employment Status and year as character to 
fill the plot but i did not find any intresting relation for that line. however 
looking at the line can see ther is nice linear relation between as line passed 
through points (400,400) and (600,600) then slowly converge to lower slope 

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
index<-which(data$IncomeRange %in% c("Not displayed" , "Not employed"))
data<-data[-index,]

ggplot(aes(x =  StatedMonthlyIncome , y=MonthlyLoanPayment ) , data = data) + 
geom_point(fill = I('#F79420') , color =I('black') , shape = 21) + labs(x = 
'monthly income' , y='monthly payment') + ggtitle("2006-2014")


```

plot here does not show much as we have high density in the middle , i can see a 
quadratic curve in the data. however i am not using log as the range of monthtly 
income and monthly payment is not that different. i will use income range and look
at the monthly payment based on the range


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
ggplot(aes(IncomeRange, 
            MonthlyLoanPayment), 
        data = subset(data, !is.na(IncomeRange)) ) +
   geom_boxplot( alpha = .5)

```

we can see as income range increases the monthly payment also increase, now i use
loan year as a factor and re plot to see if we can find any important information


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
ggplot(aes(IncomeRange, 
            MonthlyLoanPayment, fill = LoanOriginationDate.year), 
        data = subset(data, !is.na(IncomeRange)) ) +
   geom_boxplot( alpha = .5) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

we can see within last few years(2011-2014) having higher income range will be 
ending in more monthly loan payments, and intresting fact for me was that in 2009
and 2010 we are having the lowest monthly payment

# developing a Linear model

```{r echo=FALSE , message=FALSE , warning=FALSE} 

library(memisc)
require(MASS)
require(gridExtra)

m1 <- lm(MonthlyLoanPayment ~ StatedMonthlyIncome, data = data)
m2 <- update(m1, ~ . +BorrowerRate + LP_CustomerPayments + InquiriesLast6Months)
m3 <- update(m2, ~ . + LoanOriginationDate.month +  LoanOriginationDate.year + 
DebtToIncomeRatio + OpenCreditLines)
m4 <- update(m3, ~ . + LP_CustomerPrincipalPayments + LoanOriginalAmount +Term +
average_credit_score)
mtable(m1,m2,m3,m4)
```

last few plots shows that we can use some of the variables inside the model and 
add other variables and check the results, we can see we have good R-squared and
adjusted R-squared, however there is a chance of over fitting which i have not 
consider for this project. also to be sure of chosen variables are rubust i am
using an automatic variable selection based on forward and backeward regression 
using AIC metric

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
model<-lm(MonthlyLoanPayment ~ . -LoanStatus , data = data)
stepAIC(model, direction = "both")

summary(model)
```

we can see our model is really similar to automated model however we have not 
consider BankcardUtilization and EmploymentStatus which also are not statistically 
significant, also we can see variables like open credits have a negative effect of
-8.332 which kinda seems off to me as i assume having more credit lines will 
increase your monthly payment as it shows you are frequenct a borrower.


# **Final Plots and Summary**



### Plot One and Two 


we have seen Loan Origin amount has siginificant correlation with the borrower
rate, i have used employment status also to look deeper into the data and see the 
patterns, which we have seen for employment status and different 
loan amount we have differnet borrower rates.


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(14))

ggplot(aes(bucket.LoanOriginalAmount, 
            BorrowerRate, fill = EmploymentStatus), 
        data = subset(data, !is.na(bucket.LoanOriginalAmount)) ) +
   geom_boxplot( alpha = .5) + theme(axis.text.x = element_text(angle = 90, hjust = 1))+
     labs(x = "LoanOriginalAmount[$]" , y = "BorrowerRate[%]"  , 
          title ="                 borrowe rate vs loan amount ") 


```


in last plot we can see within years the average credit score has an intresting
impact to borrower rate as higher credit score will end in lower borrower rate ,
i have ploted for 2010 , 2013 and 2014 , we can see borrower rate based on credit score has decrease in recent years

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(14))

ggplot(aes(x =  average_credit_score , y=BorrowerRate) , 
data = subset(data , LoanOriginationDate.year == 2014)) + 
geom_boxplot(alpha = 0.2, color = 'blue' ,size =1)+
 ggtitle(" year 2014") + labs(x = "average_credit_score" , y = "BorrowerRate[%]",
   title = "                          borrowe rate for given credit score  ")
```


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(14))

ggplot(aes(x =  average_credit_score , y=BorrowerRate) , 
data = subset(data , LoanOriginationDate.year == 2013)) +  
geom_boxplot(alpha = 0.2, color = 'red' ,size =1)+
 ggtitle(" year 2013") + labs(x = "average_credit_score" , y = "BorrowerRate[%]",
   title = "                          borrowe rate for given credit score  ")
```


```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(14))

ggplot(aes(x =  average_credit_score , y=BorrowerRate) ,
data = subset(data , LoanOriginationDate.year == 2010)) + 
geom_boxplot(alpha = 0.2, color = 'black' ,size =1) +
ggtitle(" year 2010")  + labs(x = "average_credit_score" , y = "BorrowerRate[%]",
   title = "                          borrowe rate for given credit score  ")

```

i have found withing last few years(2011-2014) being in higher income range class
will cause to pay more payment monthly.

```{r echo=FALSE , message=FALSE , warning=FALSE} 
theme_set(theme_minimal(17))
ggplot(aes(IncomeRange, 
            MonthlyLoanPayment, fill = LoanOriginationDate.year), 
        data = subset(data, !is.na(IncomeRange)) ) +
   geom_boxplot( alpha = .5) + theme(axis.text.x = 
                                       element_text(angle = 90, hjust = 1))+
   labs(x = "Income Range [$]" , y = "Monthly loan payment[$]",
   title = "  loan payment for differnt income range")

```




#**Reflections**

## Issues 

Are the following conclusions certain? No. There seems we need more data wrangling
and cleaning on this data set.Also we need to have more records with usable values.
Regarding this project I struggled with finding the relationship between the 
variables initially. It was hard for me to decide which variables are dependent, which are
independent and which I should keep in my analysis.Through exploring, I found 
a borrower rate as my dependent variable and I designed my project around that.


## Conclusion 

My conclusions regarding this project are the following; 
The likelihood of having low/high Borrow rate depends on variables such as credit score,
amount of the loan, customer payment and other variables which model showed us,
but I did not consider them in my conclusion. 

It seems that credit score has more impact on borrow rate within last few years,
after 2010 we can see that having higher credit score gives us lower rates. 

Intrestingly I found that having higher loan amount seems to 
decrease the borrow rate, however we need more data to approve this hypothesis. 





